Water discharge has a variable impact on sediment transport capacity in subglacial systems

### 2023-09-13 Delaneyetal2023b.tex
Updates from Overleaf

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO: \color{red}#1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\mauro}[1]{{\textbf{\color{green}Mauro says:} \color{green} #1} }
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\mmms}{$\,\mathrm{m^3\, s^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

@@

    ianarburua.delaney@unil.ch},
  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland},
  Mauro A. Werder$^{3,4}$

@@

  \noindent
  In both subglacial and subaerial channels, the sediment transport capacity for sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
  In subaerial channels, changing water discharge  changes the water depth,  channel width, and velocity.
  However, in subglacial environments and on time scales of hours to weeks, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial conduit maintains a largely fixed geometry.
  Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and for subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
  Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels than subaerial ones over time and with respect to water discharge.
  Furthermore, high subglacial sediment transport capacity can persist across a wide range of water discharges.
  However, changing channel width can accommodate some  variability in sediment transport capacity in subaerial channels.
  Despite the different responses of sediment transport capacity to water discharge in the two-channel types, measurements of sediment discharge from the catchments generally correlate better with subaerial hydraulics.
  This suggests that the sediment discharge records represent other processes such as sediment access or subaerial processes in the proglacial area.
  The different responses of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, increase the occurrence of sediment transport capacity reaching mobilization thresholds, not necessarily related to water discharge quantity.

@@

In a transport-limited regime, sediment discharge responds to the sediment transport capacity defined as the amount of sediment the water can carry.
Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w_c\, v^2,
  \end{equation}
\end{linenomath*}
%
where $w_c$ is the width of the channel.
Following the conservation of mass, the velocity of the water flowing through a  channel is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q}{S},
  \end{equation}
\end{linenomath*}
where $Q$ is water discharge,  and $S$ is the channel's wetted area. %\mauro{I think ``wetted area'' is better and less ambiguous, changed throughout.}
% \mauro{later $Q$ is used.  drop $_w$?}

In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \citep{leopold1953}.  The change in water depth results in a commensurate increase in water velocity.  Thus higher discharge is accommodated by both an increase in wetted area and flow velocity such that Equation~\eqref{eq:v} remains fulfilled. As a result, the shear stress across the channel $\tau_t$ will increase according to Equation~\ref{eq:tau_t}.

The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
Therefore, subglacial water flow behaves more like pipe-flow and changes in water discharge $Q$ are mainly accommodated with increased water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.

  \begin{figure}[h]
  \centering
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different responses that subglacial and subaerial channels have to an increase in water discharge.  Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length. $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths evolve $w_{cf}$. Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed.
      Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than the depth (Section~\ref{sect:fluv}).}
    \label{fig:cartoon}
  \end{figure}
% \ian{Come back to Daniel's comments}

As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
The divergent response of changing water discharge on sediment mobilization capacity may well alter sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels, and implications for sediment discharge records, has been minimally discussed.
Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins. %\mauro{self-cite here?}

This manuscript aims to evaluate the relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
The parameterizations' lumped nature is intended to isolate the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
Furthermore, outputs are compared to sediment discharge data from these two catchments.
The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport from glacierized catchments.

@@

% \TODO{Address Daniel's comment: I think that these records need a few more words: where are they measured exactly? How long are the time series? What temporal resolution have they got? Do they cover the entire year, or only the summer months? }

@@

\section{Methods}
\label{sect:meth}
The parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
Both parameterizations calculate   water velocity, shear stress and width-integrated shear stress, upon which sediment transport depends \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \citep[e.g.][]{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.

\mauro{Yeah, I wonder whether width-integrated should be scaled by width to 2/3.}

\begin{table}[ht]
  \centering
  \caption{Variables, parameters, and constants used in this work.}
  %Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier. \mauro{those f are high}
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel wetted area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

@@

Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
The formulation here does not consider the englacial storage of water.
Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the evolution of the channel size.
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955} (usually $n=3$).
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.

Following the Darcy-Weisbach, head drop $\Delta h$ is,
\begin{linenomath*}
  \begin{equation}
    \label{eq:dh}
    \Delta h \,  = l \,\frac{1}{g}\,s\,f_r\,\frac{Q^2}{D_h^5}
  \end{equation}
\end{linenomath*}
\mauro{the $s$ is iffy}\ian{... not sure of the exact problem here... I have given the relationship between $D_h$ and $S_g$ below}
\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length over which head change occurs. $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
\begin{equation}
  \label{eq:Hf}
  s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
\end{equation}
where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
smaller values of $\beta$ result in shallow, wide channels.
The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given
\begin{equation}
  \label{eq:Dh2S}
  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta}.
\end{equation}

With knowledge of wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
\begin{equation}
  \label{eq:tau}
  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
\end{equation}
%
where  the water velocity $v_g = \frac{Q}{S_g}$.

@@

To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
the shear stress $\tau_f$ at the river bed is given by
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
Channel width $w_{cf}$ is
\begin{equation}
  \label{eq:wcf}
  w_{cf} = k \, Q^\frac{1}{2},
\end{equation}
%
where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
%\mauro{This is for so called ``regime channels''}
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
%
As in Section~\ref{sect:sub_mode}, the width-integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.

@@

The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett glacier (\icesheet).
The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
 \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).

@@

Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation, or the correlation with respect to ordering of values from the outputs and observations.
Rank correlation reduces the impact of the non-linear response of sediment discharge to the hydrological forcing and the uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.
Outputs and observations are aggregated across timescales from $1$\,\unit{hr} to $15$\,\unit{days} from the $15$\,\unit{min} outputs, by averaging values one half of the timescale before and after a point time.
%\mauro{Here a brief explanation of the rank-stuff would be good.  Also describe the time aggregation.}

@@

\subsection{Role of water discharge in sediment transport capacity}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h). Gray (black) lines in a and e represent the subglacial channel size (water discharge).  In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
    Data in plots are at $15$\unit{min} intervals.
    Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
    Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a summer.
        \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
        \ian{also, might need to go back in Timemaching to pull out old version}
      }
  \label{fig:model_outs}
\end{figure}
Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks. Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

@@

The increased performance of subaerial hydraulics, in representing observations occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}).
This could occur for two reasons.

>>>>

\newcommand{\aref}[1]{\textbf{Reference #1}}
\newcommand{\TODO}[1]{\textbf{TODO: \color{red}#1}}
\newcommand{\ian}[1]{{\textbf{\color{blue}Ian says:} \color{blue} #1} }
\newcommand{\alpine}{\textit{ALPINE}\,}
\newcommand{\icesheet}{\textit{ICESHEET}\,}
\newcommand{\m}{$\,\mathrm{m}$\,}
\newcommand{\cm}{$\,\mathrm{cm}$\,}
\newcommand{\mma}{$\,\mathrm{mm  \, a^{-1}}$\,}
\newcommand{\mmma}{$\,\mathrm{m^3\, a^{-1}}$\,}
\newcommand{\mmms}{$\,\mathrm{m^3\, s^{-1}}$\,}
\newcommand{\unit}[1]{$\mathrm{#1}$}

@@

    ianarburua.delaney@unil.ch},
  Andrew J. Tedstone\footnote{Department of Geosciences, University of Fribourg, Ch. du Musée 1700, Fribourg, Switzerland},
  Daniel Farinotti\footnote{Laboratory of Hydraulics, Hydrology and Glaciology (VAW), ETH-Z\"urich, H\"onggerbergring 26, 8093 89 Z\"urich, Switzerland}$^{,}$\footnote{Swiss Federal Institute for Forest, Snow and Landscape Research (WSL) Z\"uricherstrasse 111, 8903 1011 Birmensdorf, Switzerland}

@@

  \noindent
  In both subglacial and subaerial channels, the sediment transport capacity of a sediment size depends mainly on 1) the channel width over which the sediment is mobilized  and 2) water velocity, as this controls the shear stress that the flowing water applies to the channel bottom.
  In subaerial channels, changing water discharge  changes the water depth,  channel width, and velocity.
  However, in subglacial environments and on time scales of hours to weeks, changing water discharge  mainly alters the  water velocity, not channel area or width, because the subglacial conduit maintains a largely fixed geometry.
  Here, parameterizations of sediment transport capacity for pressurized flow under glaciers and subaerial open channel flow are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet.
  Results of these parameterizations show that sediment transport capacity varies far more in subglacial channels than subaerial ones over time and with respect to water discharge.
  Furthermore, high subglacial sediment transport capacity can persist across a wide range of water discharges.
  However, changing channel width can accommodate some  variability in sediment transport capacity in subaerial channels.
  Despite the different responses of sediment transport capacity to water discharge in the two-channel types, measurements of sediment discharge from the catchments generally correlate better with subaerial hydraulics. 
  This suggests that the sediment discharge records represent other processes such as sediment access or subaerial processes in the proglacial area.
  The different responses of water discharge on sediment transport capacity in subglacial channels, compared to subaerial ones, increase the occurrence of sediment transport capacity reaching mobilization thresholds, not necessarily related to water discharge quantity.

@@

In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment the water can carry.
Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel bottom $w_c$ over which to mobilize sediment.
The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that 
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w_c\, v^2,
  \end{equation}
\end{linenomath*}
% 
where $w_c$ is the width of the channel.
Following the conservation of mass, the velocity of the water flowing through a  channel is given as 
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q_w}{S},
  \end{equation}
\end{linenomath*}
% 
where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.

In subaerial channels, with open channel flow, the cross-sectional area $S$ of the channel evolves with changing water discharge $Q_w$  by changing both the channel's width and the water depth \citep{leopold1953}.
During a flood with greater water discharge $Q_w$, for instance, the cross-sectional area $S$ increases as the channel widens and deepens, along with water velocity $v$, which is squared. As a result, the shear stress across the channel $\tau_t$ will increase according to Equation~\ref{eq:tau_t}.

Yet, the response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
As a result, the subglacial channel size only evolves over days to months, whereas water discharge can vary over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$ \citep[e.g.][]{swift2005}, because the size of the channel $S$ changes on a much shorter timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
It follows that an increase in water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w_c$ evolves as well.

  \begin{figure}[h]
  \centering
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different responses that subglacial and subaerial channels have to an increase in water discharge.  Water velocity magnitude in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain unchanged, while the subaerial channel widths evolve $w_{cf}$. Shear stress $\tau$ is responsible for the mobilization of sediment and increases with the number of makers at the channel bed. 
      Note that in the subaerial channel parameterization, channels have rectangular shapes (Section~\ref{sect:fluv}).} 
    \label{fig:cartoon}
  \end{figure}
% \ian{Come back to Daniel's comments}

As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
These differences have been implicitly included in a wide range of available models that  quantify sediment transport from catchments in both subglacial and subaerial channels  \citep[e.g.][]{walder1994,tucker1997,creyts2013,wickert2019,hewitt2019}.
The divergent response of changing water discharge on sediment mobilization capacity may well alter sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \citep[e.g.][]{lane2016,perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may impact the interpretation of records of sediment transport in glacial systems \citep[e.g.][]{muller1968,richards2003,swift2005,ganti2016}.
Yet, the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels, and implications for sediment discharge records, has been minimally discussed.
Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins.

This manuscript aims to evaluate the relationship between sediment transport capacity and water discharge in subglacial channels, compared to subaerial ones.
Without continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher), and  a land-terminating glacier in Greenland (Leverett glacier).
The parameterizations' lumped nature is intended to isolate the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels. 
Furthermore, outputs are compared to sediment discharge data from these two catchments.
The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport from glacierized catchments.

@@

% \TODO{Address Daniel's comment: I think these records need a few more words: where are they measured exactly? How long are the time series? What temporal resolution have they got? Do they cover the entire year, or only the summer months? }

@@

\section{Methods}
\label{sect:meth}
Parameterizations below (Sections ~\ref{sect:sub_mode}~and~\ref{sect:fluv}) represent relationships amongst water discharge, water velocity, and channel geometry in both subaerial and subglacial channels (Table \ref{table:vpm}).
Both parameterizations calculate   water velocity, shear stress and width-integrated shear stress, upon which sediment transport depends \citep[Figure \ref{fig:cartoon}; ][]{shields1936}.
The evaluation of these variables omits the selection of a sediment transport relationship and grain-size parameter \citep[e.g.][]{shields1936,meyer1948} needed to calculate sediment transport capacity from shear stress.

\begin{table}[ht]
  \centering
  \caption{Variables, parameters, and constants used in this work.} 
  %Where two values are given, the first refers to  \alpine{}, a case from Fieschergletscher, and the second to a glacier marginal to the \icesheet{} a case from Leverett Glacier }
  \begin{tabular}{ l  c  c c }
    Name &Symbol&  Value&Units \\ \hline
    \textbf{Variables}  & & & \\
    Water discharge  & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
    Water velocity (glacier, subaerial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
    Channel cross-sectional area (glacier, subaerial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
    Hydraulic diameter &$D_h$&&$\mathrm{m}$\\
    Width of channel floor (glacial, subaerial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
    Hydraulic head &$\Delta h$&& $\mathrm{m}$\\
    Shear stress (glacier, subaerial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
    Width-integrated shear stress (glacier, subaerial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\

@@

Here, it is assumed that the water is transported through a subglacial channel flowing down the glacier; \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \citep[e.g.][]{rothlisberger1972,clarke2003}.
The formulation here does not consider the englacial storage of water.
Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the evolution of the channel size. 
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 (h_{o}-\frac{\Delta h}{2})^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \citep{glen1955}.
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.

Following the Darcy-Weisbach, head drop $\Delta h$ is,
\begin{linenomath*}
  \begin{equation}
    \label{eq:dh}
    \Delta h \,  = l \,\frac{1}{g}\,s\,f_r\,\frac{Q^2}{D_h^5}
  \end{equation}
\end{linenomath*}
\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length overwhich head change occurs. $s$ is a factor accounting for channel geometry \citep{hooke1990}, calculated as:
\begin{equation}
  \label{eq:Hf}
  s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
\end{equation}
where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
smaller values of $\beta$ result in shallow, wide channels.

With knowledge of cross-sectional area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
\begin{equation}
  \label{eq:tau}
  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
\end{equation}
% 
where  the water velocity $v_g = \frac{Q_w}{S_g}$.

@@

To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
Here, given  that mass is conserved mass, there is a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, and the Darcy-Weisbach relationship,
the shear stress $\tau_f$ at the river bed is represented as
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
Channel width $w_{cf}$ is 
\begin{equation}
  \label{eq:wcf}
  w_{cf} = k \, Q_w^\frac{1}{2},
\end{equation}
% 
where $k$ is a constant and the exponent is commonly given as $\frac{1}{2}$ \citep{leopold1953}.
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
% 
As in Section~\ref{sect:sub_mode}, the width-integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.

@@

The parameterizations above are applied to hydrological records from the Fieschergletscher in the Swiss Alps (\alpine) and the Leverett glacier on the Greenland Ice Sheet (\icesheet).
The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow, as it moves through the catchment.
Outputs of the parameterizations are meant to represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($<\,20$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
 \icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge (generally $<\,500$\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).

@@

Lastly,  outputs are compared with the record of sediment discharge from the two glaciers using Spearman rank correlation to reduce the impact of the non-linear response of sediment discharge to the hydrological forcing.
Rank correlation also reduces the impact uncertainties from sensor disturbances in the sediment transport record from \icesheet{}.

@@

\subsection{Role of water discharge in sediment transport capacity}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h). Gray (black) lines in a and e represent the subglacial channel size (water discharge).  In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
    Data in plots are at $15$\unit{min} intervals.
    Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
    Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a summer.
    %    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
  } 
  \label{fig:model_outs}
\end{figure}
Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks. Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

@@

Outputs from the subaerial system rank correlate higher with sediment discharge in the \alpine and \icesheet systems here.
The increased performance occurs despite  output variables showing the disparate relationship between sediment transport capacity and water discharge in subglacial systems (Figure~\ref{fig:Qw_vari}), which could occur for two reasons.

First, these measurement stations are $0.5$ to $2$\,\unit{km} downstream of the glacier terminus \citep{cowton2012,felix2022}, respectively, likely a long enough distance for the pressurized subglacial regime to transition to the open channel subaerial regime.
The transition of subglacial to subaerial conditions may occur in the transition between pressurized and subaerial flow under glacier terminus \citep{perolo2018}, especially if cavity closure rates of the ice are slow  \citep{egli2021b}.
Furthermore,  if a transport-limited regime persists at the glacier margin, then sediment can be easily mobilized and subaerial processes could be represented just a short distance downstream.
